---
title: "Analysis"
author: "Danielle Ethier"
date: "25/01/2021"
output: html_document
---

Following methods from Dunn et al. 2004

##Table 1
To determine whether banding and census had independing effects of DET we preform a multiple regression. 
DET annual index as response, and the census and banding index as the depenend variables
This will be done for each station/seaon seperatley only uing the past 10-years of data

```{r index regression}

out.dir <- paste("C:/Users/dethier/Documents/ethier-scripts/CMMN-CountComparison/output/")

#create a table for results

#create new
indexcomp<- as.data.frame(matrix(data = NA, nrow = 1, ncol = 21, byrow = FALSE, dimnames = NULL))

names(indexcomp) <- c("site", "season", "SpeciesCode", "english_name", "scientific_name", "analysis_code", "sort", "BandCensuscor1", "sig1", "DETBCcor2", "sig2", "Censussig", "Bandingsig", "R2", "Csig", "CR2", "Bsig", "BR2", "BCsig", "BCR2")

write.table(indexcomp, file = paste(out.dir,"Annual Index Comparision_June2021.csv"), row.names = FALSE, append = FALSE, quote = FALSE, sep = ",")

#Import desired inde data
index_dat<-read.csv("index_dat.csv")
index_dat<-index_dat %>% filter(kld<=5) #remove unreliable index values (i.e., outliers)

#Import sort oder 
sort<-read.csv("sp.names.csv")
sort<-sort %>% select(species_code, sort_order)
index_dat<-merge(index_dat, sort, by.x="SpeciesCode", by.y="species_code")

#Import response variable files
fallres<-read.csv("sp.site.trends.fall.csv")
fallres$season<-"Fall"
springres<-read.csv("sp.site.trends.spring.csv")
springres$season<-"Spring"
res<-rbind(fallres, springres)
res<-res %>% select(-X)

#Import trends file for comparison with res
trend<-read.csv("trend_dat.csv")
trend<-trend %>% select(SpeciesCode, site, season, analysis_code)
trends<-distinct(trend)



#Visual data check for outliers
#plot(index_dat$year, index_dat$index)

site.list<-unique(index_dat$site)
season.list<-unique(index_dat$season)

##LOOP THROUGH SITE, SEASON, SPECIES##

#Create a site specific loop 
for(m in 1:length(site.list)) {
  
#m<-1 #for testing
    
  site.data <-NULL 
  site.data <- filter(index_dat, site == site.list[m]) %>%
      droplevels()
  

#Create a season specific loop 
for(l in 1:length(season.list)) {
  
#l<-1 #for testing
    
  season.data <-NULL 
  season.data <- filter(site.data, season == season.list[l]) %>%
      droplevels()

#create specie list
species.list<-unique(season.data$SpeciesCode)  
  
#Create a species specific loop 
for(p in 1:length(species.list)) {

#p<-10 #for testing
    
  sp.data <-NULL 
  sp.data <- filter(season.data, SpeciesCode == species.list[p]) %>%
      droplevels()
  
#Create the site, season, species specific data frame for the analysis
dat<-NULL
dat<-cast(sp.data,SpeciesCode+english_name+scientific_name+sort_order+analysis_code+year~Response, value="index")
  
#correlation between Banding & Census
cor1<-NULL
try(cor1<-cor.test(dat$Band, dat$Census, method="pearson"), silent=TRUE) 

#correlation between DET & CensusBand  
cor2<-NULL
try(cor2<-cor.test(dat$DET, dat$BC, method="pearson"), silent=TRUE)    
  
#multiple regression between DET~Band+Census  
lm<-NULL
try(lm<-lm(DET~Band+Census, data=dat), silent=TRUE)

#single regression between DET~Census  
lm1<-NULL
try(lm1<-lm(DET~Census, data=dat), silent=TRUE)

#single regression between DET~Band
lm2<-NULL
try(lm2<-lm(DET~Band, data=dat), silent=TRUE)

#single regression between DET~BC  
lm3<-NULL
try(lm3<-lm(DET~BC, data=dat), silent=TRUE)

indexcomp<-NULL #clean old

#create new
indexcomp<- as.data.frame(matrix(data = NA, nrow = 1, ncol = 21, byrow = FALSE, dimnames = NULL))

names(indexcomp) <- c("site", "season", "SpeciesCode", "english_name", "scientific_name", "analysis_code", "sort", "BandCensuscor1", "sig1", "DETBCcor2", "sig2", "Censussig", "Bandingsig", "R2", "Csig", "CR2", "Bsig", "BR2", "BCsig", "BCR2")

indexcomp$site<-site.list[m]
indexcomp$season<-season.list[l]
indexcomp$SpeciesCode<-dat$SpeciesCode[1]
indexcomp$english_name<- dat$english_name[1]
indexcomp$scientific_name<-dat$scientific_name[1]
indexcomp$analysis_code<-dat$analysis_code[1]
indexcomp$sort<-dat$sort_order[1]

if (is.null(cor1)){
indexcomp$BandCensuscor1<-"NA"
indexcomp$sig1<-"NA"
} else {
indexcomp$BandCensuscor1<-cor1$estimate
indexcomp$sig1<-cor1$p.value
}

if (is.null(cor2)){
indexcomp$DETBCcor2<-"NA"
indexcomp$sig2<-"NA"
} else {
indexcomp$DETBCcor2<-cor2$estimate
indexcomp$sig2<-cor2$p.value
}

#lm=DET~Band+Census
if(is.null(lm)){
indexcomp$Censussig<-"NA"
indexcomp$Bandingsig<-"NA"
indexcomp$R2<-"NA"
}else{
try(indexcomp$Censussig<-coef(summary(lm))[3,4], silent=TRUE)
try(indexcomp$Bandingsig<-coef(summary(lm))[2,4], silent=TRUE)
try(indexcomp$R2<-summary(lm)$r.squared, silent = TRUE)
}

#lm2=DET~Band
if(is.null(lm2)){
indexcomp$Bsig<-"NA"
indexcomp$BR2<-"NA"
}else{
try(indexcomp$Bsig<-coef(summary(lm2))[2,4], silent=TRUE)
try(indexcomp$BR2<-summary(lm2)$r.squared, silent=TRUE)
}

#lm1 = DET~Census
if(is.null(lm1)){
indexcomp$Csig<-"NA"
indexcomp$CR2<-"NA"
}else{
try(indexcomp$Csig<-coef(summary(lm1))[2,4], silent=TRUE)
try(indexcomp$CR2<-summary(lm1)$r.squared, silent=TRUE)
}


#lm3=DET~BC
if(is.null(lm3)){
indexcomp$BCsig<-"NA"
indexcomp$BCR2<-"NA"
}else{
try(indexcomp$BCsig<-coef(summary(lm3))[2,4], silent=TRUE)
try(indexcomp$BCR2<-summary(lm3)$r.squared, silent=TRUE)
}

#Print final error table to file
write.table(indexcomp, file = paste(out.dir,"Annual Index Comparision_June2021.csv"), row.names = FALSE, col.name = FALSE, append = TRUE, quote = FALSE, sep = ",")


} # end species loop


   } # end season loop
    } # end site loop

```


```{r clean table}

tab <- read_csv("output/ Annual Index Comparision_June2021.csv")
tab <- tab %>% select(-X21)

tab <- tab %>% mutate(cor1sig = case_when(sig1<0.1 ~ "."), 
                      sig1<0.05 ~ "*", 
                      sig1<0.01 ~ "**", 
                      sig1<0.001 ~ "***", 
                      sig1>0.1 ~ "n.s.")

tab <- tab %>% mutate(cor2sig = case_when(sig2<0.1 ~ "."),
                      sig2<0.05 ~ "*", 
                      sig2<0.01 ~ "**", 
                      sig2<0.001 ~ "***", 
                      sig2>0.1 ~ "n.s.")

tab <- tab %>% mutate(Census = case_when(Censussig<0.1 ~ "."),
                      Censussig<0.05 ~ "*", 
                      Censussig<0.01 ~ "**", 
                      Censussig<0.001 ~ "***", 
                      Censussig>0.1 ~ "n.s.")

tab <- tab %>% mutate(Band = case_when(Bandingsig<0.1 ~ "."),
                      Bandingsig<0.05 ~ "*", 
                      Bandingsig<0.01 ~ "**", 
                      Bandingsig<0.001 ~ "***", 
                      Bandingsig>0.1 ~ "n.s.")

tab <- tab %>% mutate(BC = case_when(BCsig<0.1 ~ "."),
                      BCsig<0.05 ~ "*", 
                      BCsig<0.01 ~ "**", 
                      BCsig<0.001 ~ "***", 
                      BCsig>0.1 ~ "n.s.")

tab <- tab %>% mutate(C = case_when(Csig<0.1 ~ "."),
                      Csig<0.05 ~ "*", 
                      Csig<0.01 ~ "**", 
                      Csig<0.001 ~ "***", 
                      Csig>0.1 ~ "n.s.")

tab <- tab %>% mutate(B = case_when(Bsig<0.1 ~ "."),
                      Bsig<0.05 ~ "*", 
                      Bsig<0.01 ~ "**", 
                      Bsig<0.001 ~ "***", 
                      Bsig>0.1 ~ "n.s.")

tab<-tab %>% select(site, season, SpeciesCode, english_name, scientific_name, analysis_code, sort, BandCensuscor1, cor1sig, DETBCcor2, cor2sig, Census, Band, R2, C, CR2, B, BR2, BC, BCR2)

tab2<-merge(tab, res, by=c("SpeciesCode", "site", "season", "analysis_code"))

write.table(tab2, file = paste(out.dir,"Annual Index Comparision_Sig.csv"), row.names = FALSE, col.name = TRUE, append = TRUE, quote = FALSE, sep = ",")

```

##Table 3
Reduced major axis regression 

Compare trends from pairs of count methods within stations.
An intercept >0 indicates a tendacy to a positive bias in the first count method in each pair. If slope is < 1 it indicaes that the positive bias was less in speices with increasing trends, ect. 


```{r trend regression include=FALSE, echo=FALSE}

#Import desired data summary
trend<-read.csv("fulltrend.datasummary.csv")

#create output table
trend.regression.txt <- as.data.frame(matrix(data = NA, nrow = 1, ncol = 11, 	byrow = FALSE, dimnames = NULL))

names(trend.regression.txt) <-  c("Model", "Site", "Season", "INT", "INT_LCI", "INT_UCI", "SLOPE", "SLOPE_LCI", "SLOPE_UCI", "PVAL", "RSQ") 

write.table(trend.regression.txt, file = paste(out.dir, "trend.regression.csv", sep = ""), row.names = FALSE, append = FALSE, quote = FALSE, sep = ",")

#list for loops
season.list<-unique(trend$season)
station.list<-unique(trend$site)

#loop through site and season to generate output 

  for (m in 1:length(season.list)){
  
    #m<-1 #for testing

season.data<-NULL # clear previous  
season.data <- filter(trend, season == season.list[m]) %>% droplevels()
  
  for (n in 1:length(station.list)){
  
    #n<-1 #for testing 
    
ss.data <-NULL # clear previous    
ss.data <- filter(season.data, site == station.list[n]) %>% droplevels() 

mod1<-lmodel2(DET~Band,data=ss.data,"interval","interval",99)
mod2<-lmodel2(DET~Census,data=ss.data,"interval","interval",99)
mod3<-lmodel2(DET~BandCensus,data=ss.data,"interval","interval",99)
mod4<-lmodel2(Census~Band,data=ss.data,"interval","interval",99)

#mod1$regression.results #to print results

# wanted results from the RMA model
INT <- mod1$regression.results[[2]][4]
INT_LCI <- mod1$confidence.intervals[[2]][4]
INT_UCI <- mod1$confidence.intervals[[3]][4]
SLOPE <- mod1$regression.results[[3]][4]
SLOPE_LCI <- mod1$confidence.intervals[[4]][4]
SLOPE_UCI <- mod1$confidence.intervals[[5]][4]
PVAL <- mod1$P.param
RSQ <- mod1$rsquare
Model<-"DET vs. Band"
Site<-station.list[n]
Season<-season.list[m]
DET.band<-NULL
DET.band = data.frame(cbind(Model, Site, Season,  INT, INT_LCI, INT_UCI, SLOPE, SLOPE_LCI, SLOPE_UCI, PVAL, RSQ))

INT <- mod2$regression.results[[2]][4]
INT_LCI <- mod2$confidence.intervals[[2]][4]
INT_UCI <- mod2$confidence.intervals[[3]][4]
SLOPE <- mod2$regression.results[[3]][4]
SLOPE_LCI <- mod2$confidence.intervals[[4]][4]
SLOPE_UCI <- mod2$confidence.intervals[[5]][4]
PVAL <- mod2$P.param
RSQ <- mod2$rsquare
Model<-"DET vs. Census"
Site<-station.list[n]
Season<-season.list[m]
DET.census<-NULL
DET.census = data.frame(cbind(Model, Site, Season, INT, INT_LCI, INT_UCI, SLOPE, SLOPE_LCI, SLOPE_UCI, PVAL, RSQ))

INT <- mod3$regression.results[[2]][4]
INT_LCI <- mod3$confidence.intervals[[2]][4]
INT_UCI <- mod3$confidence.intervals[[3]][4]
SLOPE <- mod3$regression.results[[3]][4]
SLOPE_LCI <- mod3$confidence.intervals[[4]][4]
SLOPE_UCI <- mod3$confidence.intervals[[5]][4]
PVAL <- mod3$P.param
RSQ <- mod3$rsquare
Model<-"DET vs. Band+Census"
Site<-station.list[n]
Season<-season.list[m]
DET.bandcensus<-NULL
DET.bandcensus = data.frame(cbind(Model, Site, Season, INT, INT_LCI, INT_UCI, SLOPE, SLOPE_LCI, SLOPE_UCI, PVAL, RSQ))

INT <- mod4$regression.results[[2]][4]
INT_LCI <- mod4$confidence.intervals[[2]][4]
INT_UCI <- mod4$confidence.intervals[[3]][4]
SLOPE <- mod4$regression.results[[3]][4]
SLOPE_LCI <- mod4$confidence.intervals[[4]][4]
SLOPE_UCI <- mod4$confidence.intervals[[5]][4]
PVAL <- mod4$P.param
RSQ <- mod4$rsquare
Model<-"Band vs. Census"
Site<-station.list[n]
Season<-season.list[m]
band.census<-NULL
band.census = data.frame(cbind(Model, Site, Season, INT, INT_LCI, INT_UCI, SLOPE, SLOPE_LCI, SLOPE_UCI, PVAL, RSQ))

trend.regression<-NULL
trend.regression<-rbind(DET.band, DET.census, DET.bandcensus, band.census)
trend.regression$INT<-as.numeric(trend.regression$INT, round=3)
trend.regression$INT_LCI<-as.numeric(trend.regression$INT_LCI, round=3)
trend.regression$INT_UCI<-as.numeric(trend.regression$INT_UCI, round=3)
trend.regression$SLOPE<-as.numeric(trend.regression$SLOPE, round=3)
trend.regression$SLOPE_LCI<-as.numeric(trend.regression$SLOPE_LCI, round=3)
trend.regression$SLOPE_UCI<-as.numeric(trend.regression$SLOPE_UCI, round=3)
trend.regression$PVAL<-as.numeric(trend.regression$PVAL, round=3)
trend.regression$RSQ<-as.numeric(trend.regression$RSQ, round=3)

#print result to tables
write.table(trend.regression, file = paste(out.dir, "trend.regression.csv", sep = ""), row.names = FALSE, append = TRUE, quote = FALSE, sep = ",", col.names = FALSE)
	
  
} #end season loop
 } #end site loop

```



##NOT USED##

Does analysis_code or season influence whether of not we get banding, census or band+census trend output? 
Try a bernoulli/binomail glm on the output of the trend model. 0 = no trend 1 = trend 

```{r summary stats, include=FALSA, echo=FALSE}

detach(package:plyr)

fall.data <-read.csv("falltrend.datasummary.csv")
fall.data$season<-"Fall"
spring.data <-read.csv("springtrend.datasummary.csv")
spring.data$season<-"Spring"

fall.sumstat<-fall.data %>% group_by(site) %>% summarise(Tot.Species=sum(DET), Num.Band=sum(Band), Num.Census=sum(Census), Num.BC=sum(BandCensus))

spring.sumstat<-spring.data %>% group_by(site) %>% summarise(Tot.Species=sum(DET), Num.Band=sum(Band), Num.Census=sum(Census), Num.BC=sum(BandCensus))

fall.migstat<-fall.data %>% group_by(analysis_code) %>% summarise(Tot.Species=sum(DET), Num.Band=sum(Band), Num.Census=sum(Census), Num.BC=sum(BandCensus))

spring.migstat<-spring.data %>% group_by(analysis_code) %>% summarise(Tot.Species=sum(DET), Num.Band=sum(Band), Num.Census=sum(Census), Num.BC=sum(BandCensus))

```

```{r binomialcompare regression include=FALSE, echo=FALSE}

fall.data$SpeciesCode<-as.factor(fall.data$SpeciesCode)
fall.data$site<-as.factor(fall.data$site)
fall.data$analysis_code<-as.factor(fall.data$analysis_code)
fall.data$season<-as.factor(fall.data$season)

modBand<-inla(Band ~ analysis_code + f(SpeciesCode, model="iid") + f(site, model="iid"), family="binomial", data=fall.data, verbose = TRUE)
summary(modBand)



#using fall and spring combined
trend.data<-rbind(fall.data, spring.data)

trend.data$SpeciesCode<-as.factor(trend.data$SpeciesCode)
trend.data$site<-as.factor(trend.data$site)
trend.data$analysis_code<-as.factor(trend.data$analysis_code)
trend.data$season<-as.factor(trend.data$season)

modBand<-inla(Band ~ analysis_code + season + f(SpeciesCode, model="iid") + f(site, model="iid"), family="binomial", data=trend.data, verbose = TRUE)
summary(modBand)

```


