---
title: "Analysis"
author: "Danielle Ethier"
date: "25/01/2021"
output: html_document
---

Following methods from Dunn et al. 2004

##Table 1
To determine whether banding and census had independing effects of DET we preform a multiple regression. 
DET annual index as response, and the census and banding index as the depenend variables
This will be done for each station/seaon seperatley only uing the past 10-years of data

```{r index regression}

out.dir <- paste("C:/Users/dethier/Documents/ethier-scripts/CMMN-CountComparison/output/")

#create a table for results
indexcomp.txt <- as.data.frame(matrix(data = NA, nrow = 1, ncol = 9, byrow = FALSE, dimnames = NULL))

names(indexcomp.txt) <- c("SpeciesCode", "english_name", "scientific_name", "analysis_code", "Band-Census corr", "DET-Band+Census corr", "Census", "Banding", "R2") 

#Import desired inde data
index_dat<-read.csv("index_dat.csv")
index_dat<-index_dat %>% filter(kld<=5) #remove unreliable index values (i.e., outliers)

#Import sort oder 
sort<-read.csv("sp.names.csv")
sort<-sort %>% select(species_code, sort_order)
index_dat<-merge(index_dat, sort, by.x="SpeciesCode", by.y="species_code")

#Visual data check for outliers
#plot(index_dat$year, index_dat$index)

site.list<-unique(index_dat$site)
season.list<-unique(index_dat$season)
species.list<-unique(index_dat$SpeciesCode)

#Create a site specific loop 
for(m in 1:length(site.list)) {
  
m<-1 #for testing
    
  site.data <-NULL 
  site.data <- filter(index_dat, site == site.list[m]) %>%
      droplevels()
  

#Create a season specific loop 
for(l in 1:length(season.list)) {
  
l<-1 #for testing
    
  season.data <-NULL 
  season.data <- filter(site.data, season == season.list[l]) %>%
      droplevels()
  
#Create a species specific loop 
for(p in 1:length(species.list)) {
  
p<-1 #for testing
    
  sp.data <-NULL 
  sp.data <- filter(season.data, SpeciesCode == species.list[p]) %>%
      droplevels()
  
#Create the site, season, species specific data frame for the analysis
dat<-cast(sp.data, SpeciesCode+english_name+scientific_name+sort_order+year~Response, value="index")
  
#correlation between Banding & Census
cor1<-cor.test(dat$Band, dat$Census, data=dat)  
  
#correlation between DET & CensusBand  
cor2<-cor.test(dat$DET, dat$BandCensus, data=dat)    
  
#multiple regression between DET~Census+Band  
lm<-lm(DET~Band+Census, data=dat)
summary(lm)$r.squared

index.out <- NULL
index.out <- as.data.frame(matrix(indexcomp.txt))


write.table(indexcomp.txt, file = paste(out.dir, site.list[m], ".", season.list[l], ".Annual Index Comparision.csv", sep = ""), row.names = FALSE, append = FALSE, quote = FALSE, sep = ",")
  


} # end species loop
  } # end season loop
    } # end site loop

```



##Table 3
Reduced major axis regression 

Compare trends from pairs of count methods within stations.
An intercept >0 indicates a tendacy to a positive bias in the first count method in each pair. If slope is < 1 it indicaes that the positive bias was less in speices with increasing trends, ect. 


```{r trend regression include=FALSE, echo=FALSE}

#Import desired data summary
trend<-read.csv("fulltrend.datasummary.csv")

#create output table
trend.regression.txt <- as.data.frame(matrix(data = NA, nrow = 1, ncol = 11, 	byrow = FALSE, dimnames = NULL))

names(trend.regression.txt) <-  c("Model", "Site", "Season", "INT", "INT_LCI", "INT_UCI", "SLOPE", "SLOPE_LCI", "SLOPE_UCI", "PVAL", "RSQ") 

write.table(trend.regression.txt, file = paste(out.dir, "trend.regression.csv", sep = ""), row.names = FALSE, append = FALSE, quote = FALSE, sep = ",")

#list for loops
season.list<-unique(trend$season)
station.list<-unique(trend$site)

#loop through site and season to generate output 

  for (m in 1:length(season.list)){
  
    #m<-1 #for testing

season.data<-NULL # clear previous  
season.data <- filter(trend, season == season.list[m]) %>% droplevels()
  
  for (n in 1:length(station.list)){
  
    #n<-1 #for testing 
    
ss.data <-NULL # clear previous    
ss.data <- filter(season.data, site == station.list[n]) %>% droplevels() 

mod1<-lmodel2(DET~Band,data=ss.data,"interval","interval",99)
mod2<-lmodel2(DET~Census,data=ss.data,"interval","interval",99)
mod3<-lmodel2(DET~BandCensus,data=ss.data,"interval","interval",99)
mod4<-lmodel2(Census~Band,data=ss.data,"interval","interval",99)

#mod1$regression.results #to print results

# wanted results from the RMA model
INT <- mod1$regression.results[[2]][4]
INT_LCI <- mod1$confidence.intervals[[2]][4]
INT_UCI <- mod1$confidence.intervals[[3]][4]
SLOPE <- mod1$regression.results[[3]][4]
SLOPE_LCI <- mod1$confidence.intervals[[4]][4]
SLOPE_UCI <- mod1$confidence.intervals[[5]][4]
PVAL <- mod1$P.param
RSQ <- mod1$rsquare
Model<-"DET vs. Band"
Site<-station.list[n]
Season<-season.list[m]
DET.band<-NULL
DET.band = data.frame(cbind(Model, Site, Season,  INT, INT_LCI, INT_UCI, SLOPE, SLOPE_LCI, SLOPE_UCI, PVAL, RSQ))

INT <- mod2$regression.results[[2]][4]
INT_LCI <- mod2$confidence.intervals[[2]][4]
INT_UCI <- mod2$confidence.intervals[[3]][4]
SLOPE <- mod2$regression.results[[3]][4]
SLOPE_LCI <- mod2$confidence.intervals[[4]][4]
SLOPE_UCI <- mod2$confidence.intervals[[5]][4]
PVAL <- mod2$P.param
RSQ <- mod2$rsquare
Model<-"DET vs. Census"
Site<-station.list[n]
Season<-season.list[m]
DET.census<-NULL
DET.census = data.frame(cbind(Model, Site, Season, INT, INT_LCI, INT_UCI, SLOPE, SLOPE_LCI, SLOPE_UCI, PVAL, RSQ))

INT <- mod3$regression.results[[2]][4]
INT_LCI <- mod3$confidence.intervals[[2]][4]
INT_UCI <- mod3$confidence.intervals[[3]][4]
SLOPE <- mod3$regression.results[[3]][4]
SLOPE_LCI <- mod3$confidence.intervals[[4]][4]
SLOPE_UCI <- mod3$confidence.intervals[[5]][4]
PVAL <- mod3$P.param
RSQ <- mod3$rsquare
Model<-"DET vs. Band+Census"
Site<-station.list[n]
Season<-season.list[m]
DET.bandcensus<-NULL
DET.bandcensus = data.frame(cbind(Model, Site, Season, INT, INT_LCI, INT_UCI, SLOPE, SLOPE_LCI, SLOPE_UCI, PVAL, RSQ))

INT <- mod4$regression.results[[2]][4]
INT_LCI <- mod4$confidence.intervals[[2]][4]
INT_UCI <- mod4$confidence.intervals[[3]][4]
SLOPE <- mod4$regression.results[[3]][4]
SLOPE_LCI <- mod4$confidence.intervals[[4]][4]
SLOPE_UCI <- mod4$confidence.intervals[[5]][4]
PVAL <- mod4$P.param
RSQ <- mod4$rsquare
Model<-"Band vs. Census"
Site<-station.list[n]
Season<-season.list[m]
band.census<-NULL
band.census = data.frame(cbind(Model, Site, Season, INT, INT_LCI, INT_UCI, SLOPE, SLOPE_LCI, SLOPE_UCI, PVAL, RSQ))

trend.regression<-NULL
trend.regression<-rbind(DET.band, DET.census, DET.bandcensus, band.census)
trend.regression$INT<-as.numeric(trend.regression$INT, round=3)
trend.regression$INT_LCI<-as.numeric(trend.regression$INT_LCI, round=3)
trend.regression$INT_UCI<-as.numeric(trend.regression$INT_UCI, round=3)
trend.regression$SLOPE<-as.numeric(trend.regression$SLOPE, round=3)
trend.regression$SLOPE_LCI<-as.numeric(trend.regression$SLOPE_LCI, round=3)
trend.regression$SLOPE_UCI<-as.numeric(trend.regression$SLOPE_UCI, round=3)
trend.regression$PVAL<-as.numeric(trend.regression$PVAL, round=3)
trend.regression$RSQ<-as.numeric(trend.regression$RSQ, round=3)

#print result to tables
write.table(trend.regression, file = paste(out.dir, "trend.regression.csv", sep = ""), row.names = FALSE, append = TRUE, quote = FALSE, sep = ",", col.names = FALSE)
	
  
} #end season loop
 } #end site loop

```



##NOT USED##

Does analysis_code or season influence whether of not we get banding, census or band+census trend output? 
Try a bernoulli/binomail glm on the output of the trend model. 0 = no trend 1 = trend 

```{r summary stats, include=FALSA, echo=FALSE}

detach(package:plyr)

fall.data <-read.csv("falltrend.datasummary.csv")
fall.data$season<-"Fall"
spring.data <-read.csv("springtrend.datasummary.csv")
spring.data$season<-"Spring"

fall.sumstat<-fall.data %>% group_by(site) %>% summarise(Tot.Species=sum(DET), Num.Band=sum(Band), Num.Census=sum(Census), Num.BC=sum(BandCensus))

spring.sumstat<-spring.data %>% group_by(site) %>% summarise(Tot.Species=sum(DET), Num.Band=sum(Band), Num.Census=sum(Census), Num.BC=sum(BandCensus))

fall.migstat<-fall.data %>% group_by(analysis_code) %>% summarise(Tot.Species=sum(DET), Num.Band=sum(Band), Num.Census=sum(Census), Num.BC=sum(BandCensus))

spring.migstat<-spring.data %>% group_by(analysis_code) %>% summarise(Tot.Species=sum(DET), Num.Band=sum(Band), Num.Census=sum(Census), Num.BC=sum(BandCensus))

```

```{r binomialcompare regression include=FALSE, echo=FALSE}

fall.data$SpeciesCode<-as.factor(fall.data$SpeciesCode)
fall.data$site<-as.factor(fall.data$site)
fall.data$analysis_code<-as.factor(fall.data$analysis_code)
fall.data$season<-as.factor(fall.data$season)

modBand<-inla(Band ~ analysis_code + f(SpeciesCode, model="iid") + f(site, model="iid"), family="binomial", data=fall.data, verbose = TRUE)
summary(modBand)



#using fall and spring combined
trend.data<-rbind(fall.data, spring.data)

trend.data$SpeciesCode<-as.factor(trend.data$SpeciesCode)
trend.data$site<-as.factor(trend.data$site)
trend.data$analysis_code<-as.factor(trend.data$analysis_code)
trend.data$season<-as.factor(trend.data$season)

modBand<-inla(Band ~ analysis_code + season + f(SpeciesCode, model="iid") + f(site, model="iid"), family="binomial", data=trend.data, verbose = TRUE)
summary(modBand)

```


