---
title: "BBSDataManip"
author: "Danielle Ethier"
date: "26/01/2021"
output: html_document
---

```{r setup, include=FALSE}

library(tidyverse)

```

## BBS Data Manipulation 

At the time of writing, the BBS data were made available by Adam Smith with Environment and Climate Change Canada on Google drive: `https://drive.google.com/drive/folders/1OTqdql4oGTGqwDNWcEIHtpamrSifvp-f`. I have downloaded the 'All 2018 BBS Trend.csv' and saved it in the *data* folder as 'BBStrends.csv'. The following code selects and manipulates the BBS data for use in the count comparison analysis. 

```{r BBSreaddata, include=FALSE, echo=FALSE}

BBS<-read.csv(file="data/BBStrends.csv")

```

Select columns of interest

```{r BBSselect, echo=FALSE}

BBS<-BBS %>% select(Start_year, End_year, Region, Region_type, Strata_included, Trend, Percent_Change, Percent_Change_Q0.05, Percent_Change_Q0.95, Slope_Trend, Slope_Trend_Q0.05, Slope_Trend_Q0.95, Width_of_95_percent_Credible_Interval_Slope, Relative_Abundance, Reliability, species, bbs_num)

```

Filter rows of interest

```{r BBSselect, echo=FALSE}

BBS<-BBS %>% filter(Start_year>=2008 & End_year<=2018) %>% filter(Region_type %in% c("bcr", "prov_state"))

BBS<-BBS %>% filter(Region =="ON"| Region== "QC")

write.csv(BBS, "BBS.filter.csv")


```

