---
title: "DataManip"
author: "Danielle Ethier"
date: "18/12/2020"
output: html_document
---

This code was written to facilitate CMMN trend comparision using variour response variables: DET, Band, Census, Band+Census. The raw output used for this analysis can be found in the `data` folder. 


```{r load packages, include=FALSE, echo=FALSE}

library(tidyverse)
library(reshape)
library(ggplot2)
library(lmodel2) #Reduced major axis regression
library(INLA)

#custom function to plot multiple plots per page
source("./multiplotfunction.R")

out.dir<- paste("C:/Users/dethier/Documents/ethier-scripts/CMMN-CountComparison/output/")

```

Now we will bulk import the data 

```{r import trend data, include=FALSE, echo=FALSE}

require(plyr)

mydir<-"data/"
my.files<-list.files(path = mydir, patter ="*det.Trends.2018.csv", full.name=TRUE)
DETtrend<-ldply(my.files, read.csv)
DETtrend$Response<-"DET"
DETtrend<-DETtrend %>% select(-X)

my.files<-list.files(path = mydir, patter ="*band.Trends.2018.csv", full.name=TRUE)
Bandtrend<-ldply(my.files, read.csv)
Bandtrend$Response<-"Band"

my.files<-list.files(path = mydir, patter ="*census.Trends.2018.csv", full.name=TRUE)
Censustrend<-ldply(my.files, read.csv)
Censustrend$Response<-"Census"

my.files<-list.files(path = mydir, patter ="*bc.Trends.2018.csv", full.name=TRUE)
BCtrend<-ldply(my.files, read.csv)
BCtrend$Response<-"BandCensus"

BBStrend<-read.csv(paste0(mydir, "BBStrends.csv"))
BBStrend$Response<-"BBS"

detach(package:plyr) #detach or this will cause an issue with dplyr

```

```{r import index data, include=FALSE, echo=FALSE}

require(plyr)

mydir<-"data/"
my.files<-list.files(path = mydir, patter ="*det.AnnualIndices.2018.csv", full.name=TRUE)
DETindex<-ldply(my.files, read.csv)
DETindex$Response<-"DET"
DETindex<-DETindex %>% select(-X)

my.files<-list.files(path = mydir, patter ="*band.AnnualIndices.2018.csv", full.name=TRUE)
Bandindex<-ldply(my.files, read.csv)
Bandindex$Response<-"Band"

my.files<-list.files(path = mydir, patter ="*census.AnnualIndices.2018.csv", full.name=TRUE)
Censusindex<-ldply(my.files, read.csv)
Censusindex$Response<-"Census"

my.files<-list.files(path = mydir, patter ="*bc.AnnualIndices.2018.csv", full.name=TRUE)
BCindex<-ldply(my.files, read.csv)
BCindex$Response<-"BandCensus"

```

```{r compile, include=FALSE, echo=FALSE}

#combine cmmn output
cmmntrend<-rbind(Bandtrend, BCtrend, Censustrend, DETtrend)
cmmnindex<-rbind(Bandindex, BCindex, Censusindex, DETindex)

#select desired columns
cmmntrend<-cmmntrend %>% select(SpeciesCode, english_name, trnd, lower_ci, upper_ci, post_prob, period, years, season, analysis_code, site, Response)

cmmnindex<-cmmnindex %>% select(SpeciesCode, english_name, scientific_name, index, lower_ci, upper_ci, kld, period, year, season, analysis_code, site, Response)

#set filters
cmmntrend<- cmmntrend %>% filter(years=="2008-2018") %>%  #filter(analysis_code=="M") %>% 
  droplevels()

write.csv(cmmntrend, "trend_dat.csv")

cmmnindex<- cmmnindex %>% filter(year>=2008) %>%  #filter(analysis_code=="M") %>% 
  droplevels()

write.csv(cmmnindex, "index_dat.csv")

#remove NAs
#cmmntrend<-na.omit(cmmntrend)

```

#Create summary of speices monitored by each count method using trend output, analysis code considered


```{r sumstats anal code, include=FALSE, echo=FALSE}

trendf<-cmmntrend %>% filter(season=="Fall")
indexf<-cmmnindex %>% filter(season=="Fall")

fall.data<-cast(trendf, SpeciesCode+analysis_code+site~Response, value="trnd", fun.aggregate = length)
write.csv(fall.data, "falltrend.datasummary.csv")

trends<-cmmntrend %>% filter(season=="Spring")
indexs<-cmmnindex %>% filter(season=="Spring")

spring.data<-cast(trends, SpeciesCode+analysis_code+site~Response, value="trnd", fun.aggregate = length)
write.csv(spring.data, "springtrend.datasummary.csv")

full.data<-cast(cmmntrend, SpeciesCode+analysis_code+site+season~Response, value="trnd")
write.csv(full.data, "fulltrend.datasummary.csv")

```

#Table 1: Number of speices with computed trends using each response variable  

```{r T1 sumstats, include=FALSE, echo=FALSE}


fall.data<-cast(trendf, site~Response, value="trnd", fun.aggregate = length)
write.csv(fall.data, "Table1Fall.csv")


spring.data<-cast(trends, site~Response, value="trnd", fun.aggregate = length)
write.csv(spring.data, "Table1Spring.csv")

full.data<-cast(cmmntrend, site+season~Response, value="trnd")
write.csv(full.data, "Table1All.csv")

```


#Table 2: Number of speices of various migration types with computed trends using each response variable

```{r T2 sumstats, include=FALSE, echo=FALSE}


fall.data<-cast(trendf, analysis_code~Response, value="trnd", fun.aggregate = length)
write.csv(fall.data, "Table2Fall.csv")


spring.data<-cast(trends, analysis_code~Response, value="trnd", fun.aggregate = length)
write.csv(spring.data, "Table2Spring.csv")

full.data<-cast(cmmntrend, analysis_code+season~Response, value="trnd", fun.aggregate = length)
write.csv(full.data, "Table2All.csv")

```
